---
title: "1: Test article"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(readr)
library(ggplot2)
library(apstyle)
```

This is a mockup of the first article we would have.

```{r analysis}
ggplot(mtcars, aes(x = mpg, y = disp)) + geom_point() + theme_ap()
```
```{r analysis-2}
outside_data <- read_csv("data/bogus-data.csv")
ggplot(outside_data, aes(x = Category, y = Value, group = Category, color = Category)) + 
  geom_boxplot() + 
  theme_ap()
```

Things to note about the structure:

* Ease of presentation means we've chosen the `github_document` build option in Rmarkdown

* We wrap everything in the subfolder: so the raw data, the `.Rmd` file, the generated `.md` file, and the generated files (in `01-test-article_files`, usually images of plots) are all in the same folder.

* We aren't using packrat. This is partly because it overcomplicates things (it is my hope that we aren't doing 'full' data analysis in these, but rather more straightforward exposition).

* We link to and from the table of contents
